---
title: "HW 9 "
author: "Herschel & Pilgrim"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-pkg, message=FALSE, include=FALSE}
library(psych)
library(ggplot2)  # for plots
library(magrittr)  # for `%>%` operator
#library(here)
#library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(rstan)
rstan_options(auto_write = TRUE)  # save compiled STAN object
options(mc.cores = 2)  # use two cores
library(posterior)
library(bayesplot)
theme_set(theme_classic() +
    theme(panel.grid.major.y = element_line(color = "grey92")))
library(brms)

```

# Research Question

> Does Self-rated health (SRH) at baseline predict dementia risk measured by a latent dementia index at the conclusion of the study?

# Variables

- `base_srh`: self-rated health at baseline measured on a 1-3 Likert scale in response to the question: "How do you rate your overall health?" 1 = "good," 2 = "reasonable," 3 = "bad."
- `LDI_HARM`: Continuous latent factor that accounts for common variance in memory, cognitive ability and functional ability measures. Ranges from 1-11 in the sample, lower scores indicate poorer performance on dementia-indicative measures.

```{r beauty, include=FALSE}
health<-read.csv("hw9.csv")
```

## Variable Summary

```{r summ-var, echo=FALSE}
datasummary(srh2 + LDI_HARM ~ N + Mean + SD + Min + Max + Histogram, 
            data = health,)
```

## Model

Model:
$$
  \begin{aligned}
    LDI_i & \sim N(\mu_i, \sigma) \\
    \mu_i & = \eta_i \\
    \eta_i & = \beta_0 + \beta_1srh_i
  \end{aligned}
$$

Prior:
$$
  \begin{aligned}
    \beta_0 & \sim N(7, 3) \\
    \beta_1 & \sim N(0, 2) \\
    \sigma & \sim t^+_4(0,3)
  \end{aligned}
$$
# Run Model

We used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups). 

```{r m1, include=FALSE}
m1<-brm(LDI_HARM~srh2,
        family = gaussian("identity"),
        data = health,
        prior = prior(normal(7,3), class = "Intercept")+
                prior(normal(0,2), class = "b", coef = "srh2")+
                prior(student_t(4,0,3), class = "sigma"),
        iter = 4000)
```

## Results

As shown in the graphs below, the chains mixed well.

```{r rank-hist-m1, echo=FALSE}
mcmc_rank_hist(m1, pars = c("b_Intercept", "b_srh2", "sigma"))
mcmc_trace(m1, pars = c("b_Intercept", "b_srh2", "sigma"))
```

The following graph shows the posterior distributions of $\beta_0$, $\beta_1$, and $\sigma$. 

```{r summ-m1, echo=FALSE}
print(m1, statistic = "conf.int")
mcmc_dens(m1, pars = c("b_Intercept", "b_srh2", "sigma"))
```

The results show higher SRH scores predict lower LDI with with the posterior mean of the regression parameter as -0.84 and a 90% CI of [-1.20,-0.49]. This tells us that individuals who report a lower self-rated health score at baseline had higher dementia scores (worse dementia outcomes) at the end of the study.